Staying Hydrated: The Molecular Journey of Gaseous Sulfur Dioxide to a Water Surface

### 2012-03-28 resubmission/Article.tex
more of same

The bonding coordination of each \suldiox~was determined at each timestep of the simulations. Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, as a percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories. A first visual inspection reveals several trends. Clearly the ``SO'' coordination is the most populous at both temperatures. The second and third most populated coordinations are the ``S'' and ``SOO'', however their distributions differ between temperatures. In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally. The hotter temperature simulation shifts the distribution such that the ``S'' occurs 5\% less frequently than in the cold, and the ``SOO'' occurs nearly 10\% more often. The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``SOO'', respectively. These results for the hot (room temperature) system coincide with those of the previous single-temperature simulation study by Baer et al. at a similar temperature.\cite{Baer2010}

@@

	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations. Clearly the cold \suldiox~spends more time than the hot \suldiox, 13\% versus 3\%, respectively, completely unbound from the surface waters. This does not necessarily imply a complete desorption into the gas phase, but only a brief sojourn away from the waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis. Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures. Baer et al. also concluded that these three coordinations were the most frequent for a room temperature simulation, and specifically identified the ``SO'' and ``SOO'' as most common in their study.\cite{Baer2010}

	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (e.g. ``O'', ``OO'', etc.) represent the least frequently formed. A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters. In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, to the extent that a non-sulfur coordination is rarely formed during the course of all our simulations.

% Baer's assymetric oxygen binding
Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens. They concluded that there is asymmetric hydrogen bonding through the \suldiox-oxygens, with one oxygen binding more often than the other. This is supported by the findings here where all the double oxygen coordinations (e.g. ``OO'', ``SOO'', etc.) represent a much lower percentage of the coordinations than the single oxygen counterparts (e.g. ``O'', ``SO'', etc.). Furthermore, a triple-oxygen coordination (e.g. ``OOO'', ``SOOO'', etc.) is very rarely encountered. Three \suldiox-oxygen bonds only form if both \suldiox-oxygens are interacting with water hydrogens. Our finding that triple-oxygen coordinations rarely form complements Baer et al's conclusion about the asymmetry in the oxygen interactions.

Having established the preference for an interaction through the \suldiox-sulfur atom, we look to the right side of Figure \ref{fig:bonding-coordinations} at the double-sulfur coordinations (e.g. ``SS'', ``SSO'', ``SSOO'', etc.). From the data it is clear that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones. Adding a third oxygen atom is also unfavorable as the ``OOO'' and ``SSOOO'' together represent less than 1\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.

A picture can now be formed of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study. The \suldiox~will have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens. The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant coordination will be the ``SO'' bonding arrangement.

@@

The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in bonding coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot. In the cold temperature, the unbound, ``S'', and ``SO'' coordinations are more populated than in the hot systems. The increased temperature decreases the time spent in the unbound coordination, and causes all the coordinations to the right of ``SO'' in Figure \ref{fig:bonding-coordinations} to increase over the equivalent cold temperature populations. As shown, the cold \suldiox~spends nearly four times as much time unbound as the hot \suldiox, with most of the unbound population in the hot simulations shifting to coordinations with double-oxygen and double-sulfur bonds.

On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system. The relative decrease of these coordinations are matched in the hot surface by an increase of the ``SOO'' configuration. This speaks to a dramatic difference in the surface behavior of \suldiox~at the two temperatures. The cold \suldiox~spends nearly equal time in the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''. Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (e.g. addition of an oxygen bond from ``SO'' to ``SOO'', or removal of the bond from ``SO'' to ``S'') is equally probable, as long as the sulfur interaction with the water oxygen does not break.

@@

During the course of each simulation, the \suldiox~bonding coordination was determined and recorded for each timestep. From the coordination data we extract not only the populations of the various bonding coordinations, but also the frequencies of transitions between the different coordinations (i.e. the number of times each \suldiox~switched from one coordination type to another). With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} A and B, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one bonding coordination to another. The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently). Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequent transitions. Additionally, the transition lines are numbered to the right of each line with the number of times each transition occurred.

@@

As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} than transitions to less populated coordinations. Insights to the bonding process are made clearer from these graphs. In the cold system graph of Figure \ref{fig:coordination-transitions}A, we find that the majority of transitions are between the ``S-SO'' and ``SO-SOO'' nodes. The number of transitions within this ``S-SO-SOO'' group of bonding coordinations is an order of magnitude larger than any other transition. This indicates that while the \suldiox~is bound in any of the three most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other two coordinations in this group. Clearly, the \suldiox~is rather active and constantly forming and breaking bonds through its oxygens.

In the graph of the hot system in Figure \ref{fig:coordination-transitions}B, the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size. One very surprising result is in the transition from ``SOO-SOOO''. This transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations. This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox-oxygen. As noted earlier, it is likely that ``SOO'' coordinated \suldiox, asymmetrically binding twice through a single oxygen, is being pulled further into the water interface. It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox-oxygen. Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns to the ``SOO'' coordination.

>>>>

The bonding coordination of each \suldiox~was determined at each timestep of the simulations. Figure \ref{fig:bonding-coordinations} shows the distribution of bonding coordinations of the surface \suldiox, as a percentage of all bonding coordinations encountered for both the cold (blue) and hot (red) trajectories. A first visual inspection reveals several trends. At both temperatures the \suldiox~spends much of the time unbound without any bonding to the surface waters. The cluster of the three most populated coordinations is comprised of the ``S'', ``SO'', and ``SOO''. This grouping represents the coordinations with a single bond through the sulfur, and up to two interactions through the \suldiox-oxygens. In this coordination group the ``SO'' is the most populous, but in no way dominates the bonding types as the difference in populations is less than 5\% within each temperature. The second and third most populated bonding coordinations (excluding the unbound state) are the ``S'' and ``SOO'', however their distributions differ between temperatures. In cold simulations the ``S'' and ``SOO'' coordinations occur nearly equally. The hotter temperature simulation shifts the distribution such that the ``S'' occurs approximately 4\% less frequently than in the cold, and the ``SOO'' occurs nearly 8\% less often. The distribution of bonding coordinations in the hot temperature has a clear first and second most frequent coordination: ``SO'' and ``S'', respectively.

@@

	Several conclusions about the bonding behavior of \suldiox~to surface waters stem from this distribution of coordinations. Clearly the hot \suldiox~spends more time than the cold \suldiox, 45\% versus 23\%, respectively, completely unbound from the surface waters. This does not necessarily imply a complete desorption into the gas phase, but likely only a brief sojourn away from the surface waters, with all interactions and bond lengths longer than the cutoff criteria used for the analysis. The unbound coordination data is also supported by our recent spectroscopic experiments that resulted in increased \suldiox~adsorption at colder temperatures.\cite{Ota2011} Furthermore, the most frequently occurring bonding coordinations are ``S'',``SO'', and ``SOO'', with ``SO'' being the most populated at both temperatures. Baer et al. also concluded that these three coordinations were the most frequent for a room temperature simulation, and specifically identified the ``SO'' and ``SOO'' as most common in their study.\cite{Baer2010}

	Looking closer at the coordination types it is notable that coordinations lacking any sulfur interactions (e.g. ``O'', ``OO'', etc.) represent the least frequently formed. A bonding coordination with at least a single sulfur interaction is clearly favored over \suldiox~``oxygen-only'' bonding to waters. In our previous classical simulations of \suldiox~on water we concluded that during adsorption and throughout the interface, the \suldiox~orients so that its sulfur tends to the \wat~bulk side of the interface.\cite{Shamay2011} The coordination distributions here support the idea that binding through the sulfur is preferable, to the extent that a non-sulfur coordination is not often formed during the course of all our simulations.

% Baer's assymetric oxygen binding
Baer et al performed this coordination analysis for their single-temperature study, but discriminated between \suldiox~binding through the two different oxygens. They concluded that there is asymmetric hydrogen bonding through the \suldiox-oxygens, with one oxygen binding more often than the other. This is supported by the findings here where the double oxygen coordinations (e.g. ``OO'', ``SOO'', etc.) represent a lower percentage of the coordinations than the single oxygen counterparts (e.g. ``O'', ``SO'', etc.). Furthermore, a triple-oxygen coordination (e.g. ``OOO'', ``SOOO'', etc.) is very rarely encountered. Three \suldiox-oxygen bonds only form if both \suldiox-oxygens are interacting with water hydrogens. Our finding that triple-oxygen coordinations rarely form, and that double-oxygen coordination are less frequent than the single-oxygen counterparts, complements Baer et al's conclusion about the asymmetry in the oxygen interactions.

Having established the preference for an interaction through the \suldiox-sulfur atom, we look to the right side of Figure \ref{fig:bonding-coordinations} at the double-sulfur coordinations (e.g. ``SS'', ``SSO'', ``SSOO'', etc.). From the data it is clear that single-sulfur coordinations are overwhelmingly preferred over double-sulfur ones. Adding a third oxygen atom is also unfavorable as the ``OOO'', ``SOOO'', and ``SSOOO'' together represent less than 3\% of the trajectories, and the comparison between ``SOO'' to ``SOOO'' shows a very large decrease in occurrences.

A picture can now be formed of a typical \suldiox~molecule adsorbed to a water surface across both temperatures in this study. Often the \suldiox~will remain unbound from the interface, nearly twice as often when the temperature is hot. The \suldiox~will mostly have at least one interaction to neighboring waters through the sulfur, and will then bond asymmetrically through one of the oxygens either once or twice to water hydrogens. The \suldiox-oxygen bonds will form and break repeatedly throughout a trajectory, and overall the most dominant bonding coordination will be the ``SO'' bonding arrangement.

@@

The binding behavior of the \suldiox~is altered by changing the temperature of the system, as evidenced in the shift in bonding coordination populations of Figure \ref{fig:bonding-coordinations} from cold to hot. In the cold temperature, the bound coordinations are overall more populated than when hot. Conversely, the hot temperature system is much more often unbound (nearly half of the simulated time) than when the temperature is cold. As shown, the cold \suldiox~spends nearly half as much time unbound as the hot \suldiox, with most of the unbound population in the cold simulations shifting to coordinations with single-sulfur bonds.

%On the cold water surface, the ``S'' and ``SO'' are more populated than for the hot system. The relative decrease of these coordinations are matched in the hot surface by an increase of the ``SOO'' configuration. This speaks to a dramatic difference in the surface behavior of \suldiox~at the two temperatures. The cold \suldiox~spends nearly equal time in the ``S'' and ``SOO'' coordinations, but nearly 20\% more time in the ``SO''. Thus, the addition or removal of a bond through the \suldiox-oxygen to a neighboring coordination (e.g. addition of an oxygen bond from ``SO'' to ``SOO'', or removal of the bond from ``SO'' to ``S'') is equally probable, as long as the sulfur interaction with the water oxygen does not break.

@@

During the course of each simulation, the \suldiox~bonding coordination was determined and recorded for each timestep. From the coordination data we extract not only the populations of the various bonding coordinations, but also the frequencies of transitions between the different coordinations (i.e. the number of times each \suldiox~switched from one coordination type to another). With this data we have generated the directed graphs of Figure \ref{fig:coordination-transitions} depicting the cold and hot (Figure \ref{fig:coordination-transitions} left and right, respectively) bonding coordinations as circular colored nodes.\cite{Ellson2004,Gansner1999} The transitions between the coordinations are depicted as directed edges pointing in the direction of the transition from one bonding coordination to another. The populations of the coordinations are depicted by both the node size and coloration (larger and darker red coordinations occur more frequently). Populations of the transitions between coordinations are depicted by arrow thickness, with thicker lines corresponding to more frequent transitions. Additionally, the transition lines are numbered to the right of each line with the number of times each transition occurred.

@@

As expected for the more populated coordinations, there are more transitions between larger nodes in Figure \ref{fig:coordination-transitions} (left) than transitions to less populated coordinations. Insights to the bonding process are made clearer from these graphs. In the cold system graph of Figure \ref{fig:coordination-transitions}, we find that the majority of transitions are between the unbound and ``O'' nodes, and between the single-sulfur coordinations (e.g. ``S-SO'', ``SO-SOO'', etc.) The number of transitions within this group of bonding coordinations is an order of magnitude larger than any other transition. This indicates that while the \suldiox~is bound in any of the most populated bonding coordinations, it is actively binding and unbinding the oxygens to form the other coordinations. Clearly, the \suldiox~is rather active and constantly forming and breaking bonds through its oxygens.

In the graph of the hot system in Figure \ref{fig:coordination-transitions} (right), the transition frequencies follow the same trend as in the cold system, increasing with adjacent node size. One very surprising result is in the transition from ``SOO-SOOO''. This transition frequency does not follow from the adjacent node sizes, as the ``SOOO'' node represents less than 5\% of the bonding coordinations. This is indicative of a very rapid cycle of forming and breaking of bonds to the second \suldiox-oxygen. As noted earlier, it is likely that ``SOO'' coordinated \suldiox, asymmetrically binding twice through a single oxygen, is being pulled further into the water interface. It is likely more surrounded by waters, and in the hot system it can more easily form a brief third hydrogen bond to a water through the second \suldiox-oxygen. Because the triple-oxygen coordination is not as favorable, it quickly breaks the bond and the \suldiox~returns to the ``SOO'' coordination.
